Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-09-27 response_letter_2/response_letter.tex

`` MSMs were constructed for each of the four DNA sequences \roodr{at their respective 3SPN.2 melting temperatures at 7 mM concentration} from the 40$\times$25 $\mu$s simulation trajectories''

On page 20:

``Since each system is simulated at the \roodr{same 7 mM concentration} and at its respective melting temperature''

(III) We thank the reviewer for the suggestion to correct our simulation prediction of the melting temperature to the experimental concentration. The proposed strategy is to assume ideal behavior, reduce the computed $\phi$ ratio per Eqn.~7 in Ouldridge et al.\ \cite{Ouldridge2010ExtractingSimulations} by the 3.5$\times$ concentration difference, and then recompute $f_\mathrm{hybridized}$. We believe that this procedure would work under the assumption that the equilibrium constant does not change with temperature, but in general this may not be a good assumption since $-R T_m K_{eq} = \Delta G^\circ = \Delta H^\circ - T_m \Delta S$. Owczarzy et al.\ \cite{Owczarzy1997predicting} engage this issue to develop a van't Hoff-like concentration correction to $T_m$ based on the enthalpies of duplex formation and helix nucleation. In the absence of knowledge of these standard enthalpies, we could instead calculate the melting temperature by performing simulations over a range of temperatures, applying the proposed concentration correction to predict $f_\mathrm{hybridized}$ at that temperature, then fitting a melting curve from which we could extract the concentration-corrected $T_m$. This additional suite of calculations would be exclusively in service of nailing down the concentration correction and would not change our approach since we intentionally selected the 7 mM 3SPN.2 melting temperature in order to maximize the number of spontaneous transitions between dissociated and hybridized states to provide a rich data set of these events over which to train our MSM.

The approach we take to the concentration correction in the manuscript is an empirical one, wherein we draw a comparison of the computational predictions against experimental measurements. The observed calibration corrections therefore account for net effect of both the 3.5$\times$ higher concentration in simulation and the assumptions and approximations inherent in the 3SPN.2 model. As the reviewer observes, the (+4) K calibration correction encompassing both of these effects is overly optimistic since the concentration-corrected $T_m$ will be depressed and controlling for this would require a larger positive correction for the error due to the 3SPN.2 model approximations.
